<<<<<<< HEAD
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Set working directory.  *Be sure* to edit if reproducing.

```{r}
setwd("~/GitHub/RepData_PeerAssessment1")
```

Unzip file, read csv file, and check structure
```{r}
unzip("activity.zip")
activity<-read.csv("activity.csv")
str(activity)
```

## What is mean total number of steps taken per day?
Sum total steps for each day ignoring NAs.  Create histogram of results.
```{r total_steps_per_day_histogram, fig.keep="all", fig.show="asis"}
step_tot<-aggregate(activity$steps,by=list(date=activity$date),FUN=sum,na.rm=TRUE)
hist(step_tot$x)
```
Most days fall between 10,000 and 15,000 steps per day

Calculate mean and median steps taken per day
```{r}
mean(step_tot$x,na.rm=TRUE)
median(step_tot$x,na.rm=TRUE)
```


## What is the average daily activity pattern?
Calculate mean steps for interval period across all days.
```{r avg_daily_pattern, fig.keep="all", fig.show="asis"}
step_int<-aggregate(activity$steps,by=list(interval=activity$interval),FUN=mean,na.rm=TRUE)
plot(step_int$interval,step_int$x,xlab="Interval",ylab="Mean Steps",type="l")
```
It appears that the most active interval is around 800.  Confirm with calculating max.
```{r}
t<-max(step_int$x)
step_int[step_int$x==t,]
```
Most active interval is 8:35 am

## Imputing missing values
How often is data missing (NA) in the dataset?
```{r}
sum(is.na(activity$steps))
```

Need strategy to impute missing step values.  Propose using the mean steps for the interval.
Will use mean steps for interval and join to original data set.  They will calculate imputed steps if NA.
```{r}
step_merge<-merge(activity,step_int)
step_merge$imp<-ifelse(is.na(step_merge$steps),step_merge$x,step_merge$steps)
head(step_merge)
```

Calculate total steps for each day, to create histogram
```{r histogram2, fig.keep="all", fig.show="asis"}
step_new<-aggregate(step_merge$imp,by=list(date=activity$date),FUN=sum,na.rm=TRUE)
hist(step_new$x,xlab="Mean steps per day")
```

Calculate mean and median imputed steps
```{r}
mean(step_new$x)
median(step_new$x)
```
Both mean and median steps have increased due to this imputation logic


## Are there differences in activity patterns between weekdays and weekends?
Need to calculate weekdays after converting "date" to posix

```{r}
step_merge$days<-weekdays(strptime(step_merge$date,"%Y-%m-%d"))
table(step_merge$ days)
```
Looks like we have more data for weekdays than weekends.  Calculate new field as factor to flag weekend vs weekday.

```{r}
step_merge$type<-as.factor(ifelse(step_merge$days %in% c("Saturday","Sunday"),"Weekend","Weekday"))
table(step_merge$type)
```

Calculate total steps per day
```{r}
step_type<-aggregate(step_merge$imp,by=list(type=step_merge$type,interval=step_merge$interval),FUN=mean)
head(step_type)
```

Plot average steps per interval for weekdays vs Weekends.
```{r weekend_weekday, fig.keep="all", fig.show="asis"}
library(ggplot2)
qplot(interval,x,data=step_type,facets=type~.,ylab="Avg. Steps")+geom_line()
```

=======
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Set working directory.  *Be sure* to edit if reproducing.

```{r}
setwd("~/Coursera/Reproducible Research/RepData_PeerAssessment1-master/RepData_PeerAssessment1-master")
```

Unzip file, read csv file, and check structure
```{r}
unzip("activity.zip")
activity<-read.csv("activity.csv")
str(activity)
```

## What is mean total number of steps taken per day?
Sum total steps for each day ignoring NAs.  Create histogram of results.
```{r total_steps_per_day_histogram, fig.keep="all", fig.show="asis"}
step_tot<-aggregate(activity$steps,by=list(date=activity$date),FUN=sum,na.rm=TRUE)
hist(step_tot$x)
```
Most days fall between 10,000 and 15,000 steps per day

Calculate mean and median steps taken per day
```{r}
mean(step_tot$x,na.rm=TRUE)
median(step_tot$x,na.rm=TRUE)
```


## What is the average daily activity pattern?
Calculate mean steps for interval period across all days.
```{r avg_daily_pattern, fig.keep="all", fig.show="asis"}
step_int<-aggregate(activity$steps,by=list(interval=activity$interval),FUN=mean,na.rm=TRUE)
plot(step_int$interval,step_int$x,xlab="Interval",ylab="Mean Steps",type="l")
```
It appears that the most active interval is around 800.  Confirm with calculating max.
```{r}
t<-max(step_int$x)
step_int[step_int$x==t,]
```
Most active interval is 8:35 am

## Imputing missing values
How often is data missing (NA) in the dataset?
```{r}
sum(is.na(activity$steps))
```

Need strategy to impute missing step values.  Propose using the mean steps for the interval.
Will use mean steps for interval and join to original data set.  They will calculate imputed steps if NA.
```{r}
step_merge<-merge(activity,step_int)
step_merge$imp<-ifelse(is.na(step_merge$steps),step_merge$x,step_merge$steps)
head(step_merge)
```

Calculate total steps for each day, to create histogram
```{r histogram2, fig.keep="all", fig.show="asis"}
step_new<-aggregate(step_merge$imp,by=list(date=activity$date),FUN=sum,na.rm=TRUE)
hist(step_new$x,xlab="Mean steps per day")
```

Calculate mean and median imputed steps
```{r}
mean(step_new$x)
median(step_new$x)
```
Both mean and median steps have increased due to this imputation logic


## Are there differences in activity patterns between weekdays and weekends?
Need to calculate weekdays after converting "date" to posix

```{r}
step_merge$days<-weekdays(strptime(step_merge$date,"%Y-%m-%d"))
table(step_merge$ days)
```
Looks like we have more data for weekdays than weekends.  Calculate new field as factor to flag weekend vs weekday.

```{r}
step_merge$type<-as.factor(ifelse(step_merge$days %in% c("Saturday","Sunday"),"Weekend","Weekday"))
table(step_merge$type)
```

Calculate total steps per day
```{r}
step_type<-aggregate(step_merge$imp,by=list(type=step_merge$type,interval=step_merge$interval),FUN=mean)
head(step_type)
```

Plot average steps per interval for weekdays vs Weekends.
```{r weekend_weekday, fig.keep="all", fig.show="asis"}
library(ggplot2)
qplot(interval,x,data=step_type,facets=type~.,ylab="Avg. Steps")+geom_line()
```

>>>>>>> f94452ce4927e88635b56602751e6fe4298e13fc
